% Blueprint content for the prime_gap formalisation.
%
% Conventions used throughout this file:
%   \rocq{PrimeGapS1.<File>.<name>}  -- link to a Rocq declaration
%   \rocqok                          -- the surrounding env is fully Qed-formalised
%   \uses{label1, ...}               -- dep-graph dependencies
%   \label{...}                      -- unique label, prefixed by env type
%
% The Rocq logical name of every theory file is `PrimeGapS1.<File>` (see
% `_CoqProject`), so e.g. `CertPencil.maynard_M105_certified_pencil` becomes
% `PrimeGapS1.CertPencil.maynard_M105_certified_pencil`.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter*{Introduction}

This blueprint accompanies a Rocq formalisation that machine-verifies the
numerical step underpinning Maynard's bounded-gap theorem:
\[
   \Mk[105] \;=\; 105 \cdot \lammax\!\bigl(\Mone^{-1}\Mtwo\bigr) \;>\; 4.
\]
The headline is formula~(8.15) of James Maynard,
\emph{Small gaps between primes} (arXiv:1311.4600;
Annals of Mathematics \textbf{181} (2015), 383--413).
In the published proof, this inequality is checked by a Mathematica
notebook (\texttt{Computations.nb}, supplied as supplementary material
with the arXiv preprint). The present project replaces that
Mathematica-based computation with a Rocq proof that re-derives every
algebraic claim inside the kernel.

The end-to-end statement is
\begin{verbatim}
Definition M105 : 'M[rat]_42 := 105%:Q *: A_rat.

Definition matches_closed_forms (M : 'M[rat]_42) : Prop :=
  [/\ M = M105,
      (forall i j, (i < 42)%nat -> (j < 42)%nat ->
         M1_spec_ij i j = Z2rat (mat_get M1_int i j) / Z2rat D_M1)
    & (forall i j, (i < 42)%nat -> (j < 42)%nat ->
         M2_spec_ij i j = Z2rat (mat_get M2_int i j) / Z2rat D_M2)].

Theorem maynard_M105_certified_pencil :
  matches_closed_forms M105 /\
  exists lambda : realalg,
    eigenvalue (map_mx (ratr : rat -> realalg) M105) lambda
    /\ (4 < lambda).
\end{verbatim}
proven in \texttt{theories/S1/CertPencil.v}, where
$M_{105} := 105\cdot\Arat$ (so $M_{105} = 105\cdot\Mone^{-1}\Mtwo$) and
$\texttt{Z2rat}\,(z : \Z) : \rat := (\texttt{Z\_to\_int}\,z)\%\texttt{:\~{}R}$
is the thin embedding of $\Z$ into $\rat$. This headline is now the same
shape as the \texttt{main} branch's \texttt{maynard\_M105\_certified}
($\texttt{matches\_closed\_forms}\,M_{105}$ plus a real eigenvalue of
$M_{105}$ strictly above $4$), except it is stated over $\realalg$ where
\texttt{main} uses \texttt{algC}. The $\texttt{matches\_closed\_forms}\,M_{105}$
conjunct packages the definitional $M_{105} = 105\cdot\Arat$ together with
two composed per-matrix identities: the $\rat$-level paper-form
spec $\texttt{M\{1,2\}\_spec\_ij}$ (the readable transcription of
Maynard's Lemma~8.2) is equal, as a rational, to the FLINT-shipped
integer entry $\texttt{mat\_get}\,M_{1,2}^{\mathrm{int}}\,i\,j$ divided
by the common denominator $\DMone$ / $\DMtwo$. The composition collapses
the trust chain
$\text{FLINT integer matrices} = \text{$\Z$-level closed form}
 = \text{$\rat$-level paper-form spec}$
into one displayed equality per matrix, with the $\Z$-level boolean
checks $\texttt{all\_match\_M\{1,2\}Z} = \texttt{true}$ and the
$\rat\leftrightarrow\Z$ bridges $\texttt{M\{1,2\}\_spec\_rat\_eq}$
moved into the proof body as implementation details. (Both remain
available as standalone \Qed{}s in \texttt{MaynardVerify.v} and
\texttt{MaynardSpecBridge.v} for auditors who want to trace the chain
step by step.) Together with the eigenvalue conjunct, these certify
$\text{eigenvalue} > 4/105$ under a single \texttt{Print Assumptions},
which reports only Rocq's \texttt{PrimInt63} kernel primitives: the
project itself contains zero \texttt{Axiom} declarations and zero
\texttt{Admitted} lemmas.

\paragraph{Scope.}
What this Rocq layer proves, kernel-checked, is the left-hand side of
\[
\underbrace{\;\lammax(\Mone^{-1}\Mtwo) \;>\; 4/105\;}_{\text{Rocq, kernel-checked}}
\quad\Longrightarrow\quad
\underbrace{\;\Mk[105] = 105\cdot\lammax \;>\; 4\;}_{\text{Maynard's paper, Lemma 8.3}}.
\]
The right arrow is Maynard's Lemma~8.3 (a generalised Rayleigh-quotient
identity proved in the Annals paper, refereed there); the project does
not re-formalise this analytic step. The project's scope is exactly to
replace the Mathematica computation, not to re-formalise Maynard's paper.

\paragraph{Two distinct certificates for the same fact.}
Maynard's notebook proves $\lammax > 4/k$ via the \emph{eigenvector
route}: it numerically computes the top eigenvector of $\Mone^{-1}\Mtwo$,
snaps it to a small-denominator rational vector, and evaluates the
Rayleigh quotient $k\cdot v^T M_2 v / v^T M_1 v$ in exact rational
arithmetic. This is a true rigorous lower bound regardless of how~$v$
is obtained. Our Rocq proof takes the \emph{pencil-determinant route}:
a generic identity of commutative-ring algebra,
\[
   \det(\lambda\,M_1 - M_2)
     \;=\; \det(M_1)\,\cdot\,\chi_{M_1^{-1}M_2}(\lambda),
\]
(Rocq lemma \rocq{PrimeGapS1.DetPencil.det_pencil}, one-line
mathcomp proof), specialised at $\lambda = 4/105$ and rescaled to
integers, reduces the eigenvalue bound to the sign of two integer
determinants.  Both signs are extracted by a single shared
$710$-prime CRT lift on the \emph{constant terms} of the
characteristic polynomials, and the intermediate value theorem on
$\charpolysym_{\Arat}$ (MathComp's \texttt{poly\_ivtoo} together
with \texttt{cauchy\_bound}) certifies the existence of an
eigenvalue strictly above $4/105$.  No eigenvector is constructed.
Both routes are sound; together, they form an independent
double-check of Maynard's headline value $\Mk[105] \approx 4.0021$.

\paragraph{How to read this blueprint.}
Each chapter mirrors a Rocq source file or a small group of related
files. Inside each chapter, mathematical statements are presented as
\texttt{lemma} / \texttt{theorem} / \texttt{definition} environments,
each tagged with the corresponding Rocq declaration via
\verb|\rocq{...}|. The \verb|\rocqok| marker means the surrounding
environment is fully formalised and closed by \verb|Qed| in the source
tree. \verb|\uses{...}| declarations populate the dependency graph
shown in the web version of the blueprint.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{The matrix construction}\label{ch:matrix}

This chapter sets up the input data that the rest of the proof consumes:
the 42-element basis, the closed-form entries of $\Mone$ and $\Mtwo$,
the $\Z$-level cross-multiplication checks that anchor the
FLINT-shipped integer matrices to those closed forms, the kernel-Qed
bridge between the rat-level paper-form spec and the Z-level
computational spec, and the bridge into MathComp's $\Q$-level matrices.
All material here lives inside the files \texttt{MaynardSpec.v},
\texttt{MaynardBasis.v}, \texttt{MaynardFactQ.v}, the
\texttt{MaynardVerify/} directory (\texttt{Def.v}, six parallel
\texttt{M2\_0..5.v} chunks, and the assembly \texttt{MaynardVerify.v}),
and \texttt{MaynardSpecBridge.v}. Throughout this chapter we follow the
v3/Annals numbering of Maynard's paper: Lemma~8.1 / 8.2 / 8.3 (=
Lemma~7.1 / 7.2 / 7.3 in v1).

\section{Conventions}

\begin{convention}
The two power sums are
$P_1 = \sum_i t_i$ and $P_2 = \sum_i t_i^2$. The basis monomials are
$F_{b,c} = (1 - P_1)^b \cdot P_2^c$ with $b + 2c \le 11$. Indices
$(b_i, c_i)$, $(b_j, c_j)$ denote bidegrees of the $i$-th and $j$-th
basis monomials. The outer integration dimension is $k = 105$; we use
two constants $K_1 = 105$ (outer simplex) and $K_2 = 104 = K_1 - 1$
(residual simplex after the eq.~8.8 substitution that integrates out
$t_1$).
\end{convention}

\begin{definition}[Constants $K_1$, $K_2$]\label{def:K1K2}
\rocq{PrimeGapS1.MaynardSpec.K1, PrimeGapS1.MaynardSpec.K2}
\rocqok
$K_1 := 105$, $K_2 := 104$. The first is the outer simplex dimension
(used in $M_1$ and in the threshold $4/K_1$); the second is the
residual simplex dimension after the eq.~8.8 substitution (used in $M_2$).
\end{definition}

\section{The 42-element basis}

\begin{definition}[Basis]\label{def:maynard_basis}
\rocq{PrimeGapS1.MaynardBasis.maynard_basis}
\rocqok
The Maynard basis is the explicit list of 42 pairs $(b, c) \in \N^2$
with $b + 2c \le 11$, in the Mathematica enumeration order used by
\texttt{flint\_probe.py}:
\[
[(0,0), (1,0), (0,1), (2,0), (1,1), (3,0),\, (0,2), \ldots, (9,1), (11,0)].
\]
\end{definition}

\begin{lemma}[Basis size]\label{lem:basis_size}
\rocq{PrimeGapS1.MaynardBasis.maynard_basis_size}
\rocqok
\uses{def:maynard_basis}
$\#\{(b, c) \in \N^2 : b + 2c \le 11\} = 42$.
\end{lemma}
\begin{proof}\rocqok
$\sum_{c=0}^{5}(12 - 2c) = 12 + 10 + 8 + 6 + 4 + 2 = 42$.
A single \vmcompute{} on the list closes it.
\end{proof}

\begin{lemma}[Basis matches FLINT enumeration]\label{lem:basis_matches}
\rocq{PrimeGapS1.MaynardBasis.maynard_basis_eq_witness}
\rocqok
\uses{def:maynard_basis}
$\texttt{maynard\_basis} = \texttt{Witness.basis}$,
i.e.\ the hand-readable basis in \texttt{MaynardBasis.v} agrees
literally with the autogenerated basis used to index the shipped
integer matrices in \texttt{Witness.v}.
\end{lemma}
\begin{proof}\rocqok
\vmcompute{} on a list of 42 pairs.
\end{proof}

\begin{lemma}[Basis is the canonical set]\label{lem:basis_spec}
\rocq{PrimeGapS1.MaynardBasis.maynard_basis_spec}
\rocqok
\uses{def:maynard_basis}
For every $p \in \N \times \N$,
\[
  p \in \texttt{maynard\_basis} \quad\iff\quad p.1 + 2 \cdot p.2 \le 11.
\]
\end{lemma}
\begin{proof}\rocqok
The basis is permutation-equal to the canonical filter
\[
  \texttt{canonical\_basis} \;:=\; \bigl[\, p \,\big|\, p \in
    \texttt{iota}\,0\,12 \times \texttt{iota}\,0\,6,\; p.1 + 2\,p.2 \le 11
  \bigr]
\]
by a \vmcompute{} \texttt{perm\_eq} lemma, and \texttt{canonical\_basis}
satisfies the spec by construction (\texttt{mem\_filter} +
\texttt{allpairs\_f}; the residual $2c \le 11 \Rightarrow c < 6$ step
closes by \texttt{lia}).
\end{proof}

\begin{lemma}[Basis has no duplicates]\label{lem:basis_uniq}
\rocq{PrimeGapS1.MaynardBasis.maynard_basis_uniq}
\rocqok
\uses{def:maynard_basis}
$\texttt{uniq}\;\texttt{maynard\_basis}$.
\end{lemma}
\begin{proof}\rocqok
\vmcompute{} on the 42-pair list.
\end{proof}

\begin{remark}
Lemmas~\ref{lem:basis_size}, \ref{lem:basis_spec},
and~\ref{lem:basis_uniq} together pin \texttt{maynard\_basis} to
exactly the multiset $\{(b, c) \in \N^2 : b + 2c \le 11\}$. A reviewer
never has to inspect the literal 42-pair list — the basis is fully
characterised by the predicate $b + 2c \le 11$, the size $42$, and the
absence of duplicates. Order matters only because the matrix rows and
columns must align with the indexing of $M_{1,2}^{\mathrm{int}}$;
eigenvalues of $M_1^{-1} M_2$ are invariant under simultaneous
row/column permutation, so any ordering of the same set yields the
same $\lammax$.
\end{remark}

\begin{remark}
This is a lower-bound subspace argument. Maynard's full optimisation is
over all symmetric $F$ of degree $\le 11$ in $k = 105$ variables;
restricting to polynomials in the two power sums $P_1, P_2$ is a
strict subspace, and the supremum of a Rayleigh quotient over a
subspace is $\le$ the unrestricted supremum. So a lower bound of
$4/105$ on the 42-dimensional restricted problem yields
$\Mk[105] > 4$.
\end{remark}

\section{The polynomial \texorpdfstring{$G_{n,2}(k)$}{G\_n,2(k)} (Lemma 8.1)}

\begin{definition}[Compositions of $n$]\label{def:compositions}
\rocq{PrimeGapS1.MaynardSpec.compositions}
\rocqok
For $r, n \in \N$,
$\Comp(r, n) := \{(b_1, \ldots, b_r) \in \N_{\ge 1}^r : \sum_s b_s = n\}$,
returned as a list in lexicographic order.  For $r = 0$ this is the
singleton $\{[\,]\}$ if $n = 0$ and empty otherwise; for $r = n + 1$
or larger it is empty.
\end{definition}

\begin{definition}[Inner factor $\Cff(a)$]\label{def:cff}
\rocq{PrimeGapS1.MaynardSpec.cff}
\rocqok
\uses{def:compositions}
For a composition $a = (b_1, \ldots, b_r)$,
\[
\Cff(a) \;:=\; \prod_{s=1}^{r} \frac{(2 b_s)!}{b_s!}.
\]
This is exactly the per-composition inner product appearing in
Maynard's Lemma~8.1 (with $j = 2$, so $j b_s = 2 b_s$).
\end{definition}

\begin{definition}[Maynard's $G_{n,2}(k)$]\label{def:G_2}
\rocq{PrimeGapS1.MaynardSpec.G_2}
\rocqok
\uses{def:compositions, def:cff}
For $n, k \in \N$:
$\Gnk{0}{k} = 1$, and for $n \ge 1$,
\[
\Gnk{n}{k} \;=\; n! \cdot \sum_{r = 1}^{n} \binom{k}{r}
                          \sum_{a \in \Comp(r, n)} \Cff(a).
\]
This is Lemma~8.1 of arXiv:1311.4600~v3 character-for-character: the
$n!$ prefactor, the outer $r$-sum, the binomial coefficient, the
inner sum over length-$r$ compositions of $n$, and the inner product
$\prod_s (j b_s)!/b_s!$ with $j = 2$ all appear at the same positions.
\end{definition}

\begin{lemma}[Sanity values of $G_{n,2}$]\label{lem:G_2_small}
\rocqok
\uses{def:G_2}
$\Gnk{0}{k} = 1$, $\Gnk{1}{k} = 2k$, $\Gnk{2}{k} = 4 k^2 + 20 k$.
\end{lemma}
\begin{proof}\rocqok
$\Gnk{1}{k}$: the only length-1 composition of $1$ is $[1]$, with
$\Cff([1]) = 2!/1! = 2$, so the sum is $1! \cdot \binom{k}{1} \cdot 2 = 2k$.
$\Gnk{2}{k}$: length-1 contributes $\binom{k}{1} \cdot \Cff([2]) = k \cdot 4!/2! = 12k$;
length-2 contributes $\binom{k}{2} \cdot \Cff([1,1]) = \binom{k}{2} \cdot (2 \cdot 2) = 2k(k-1)$;
total $= 2! \cdot (12k + 2k(k-1)) = 2(12k + 2k^2 - 2k) = 4k^2 + 20k$.
Both also reduce by \vmcompute{}.
\end{proof}

\section{Closed forms for \texorpdfstring{$M_1$ and $M_2$}{M\_1 and M\_2} (Lemma 8.2)}

\begin{definition}[$M_1$ entry]\label{def:M1_entry}
\rocq{PrimeGapS1.MaynardSpec.M1_entry}
\rocqok
\uses{def:G_2, def:K1K2}
For $(b_i, c_i, b_j, c_j)$ basis bidegrees, set $b = b_i + b_j$ and
$c = c_i + c_j$. Then
\[
M_{1, ij} \;=\; \frac{b!}{(K_1 + b + 2c)!} \cdot \Gnk{c}{K_1}.
\]
This is Lemma~8.2 (first part) of arXiv:1311.4600~v3 verbatim, with
$K_1 = k = 105$.
\end{definition}

\begin{definition}[Eq.~8.8 expansion coefficient $\alpha$]\label{def:alpha}
\rocq{PrimeGapS1.MaynardSpec.alpha}
\rocqok
For $b, c, c' \in \N$ with $c' \le c$,
\[
\alpha(b, c, c') \;=\;
   \frac{\binom{c}{c'} \cdot b! \cdot (2c - 2c')!}{(b + 2c - 2c' + 1)!}.
\]
This is the coefficient of
$(1 - \sum_{j \ne 1} t_j)^{b + 2c - 2c' + 1} \cdot (\sum_{j \ne 1} t_j^2)^{c'}$
in the expansion of
$\int_0^{1 - \sum_{j \ne 1} t_j} (1 - P_1)^b P_2^c \, dt_1$,
obtained by expanding $P_2^c = (t_1^2 + \sum_{j \ne 1} t_j^2)^c$
binomially and integrating $t_1^{2(c - c')} (1 - \sum_j t_j)^b$
via the Beta-function identity
$\int_0^X t^a (X - t)^b \, dt = a! b! X^{a + b + 1} / (a + b + 1)!$.
\end{definition}

\begin{definition}[$M_2$ entry]\label{def:M2_entry}
\rocq{PrimeGapS1.MaynardSpec.M2_entry}
\rocqok
\uses{def:alpha, def:G_2, def:K1K2}
For $(b_i, c_i, b_j, c_j)$ basis bidegrees, with
$b'_1 = b_i + 2c_i - 2c'_1 + 1$, $b'_2 = b_j + 2c_j - 2c'_2 + 1$,
$b_{\mathrm{sum}} = b'_1 + b'_2$, $c_{\mathrm{sum}} = c'_1 + c'_2$:
\[
M_{2, ij}^{(1)} \;=\;
   \sum_{c'_1 = 0}^{c_i} \sum_{c'_2 = 0}^{c_j}
     \alpha(b_i, c_i, c'_1)\cdot \alpha(b_j, c_j, c'_2)
     \cdot \frac{b_{\mathrm{sum}}!}{(K_2 + b_{\mathrm{sum}} + 2 c_{\mathrm{sum}})!}
     \cdot \Gnk{c_{\mathrm{sum}}}{K_2}.
\]
This is Lemma~8.2 (second part) of arXiv:1311.4600~v3 for the
\emph{per-coordinate} integrand $J_k^{(1)}$ (not $\sum_m J_k^{(m)}$):
the symmetry factor $k$ is paid in the threshold $4/k$, not in the
matrix entries.
\end{definition}

\begin{remark}
The matrices have up to $36$ rational summands per entry (when
$c_i = c_j = 5$), with intermediate denominators reaching
$\sim 10^{7000}$ digits before the rational normalisation collapses
them. This is why \texttt{MaynardSpec.v} also ships
$\Z$-level twins (\texttt{m1\_num\_den}, \texttt{m2\_num\_den}) that
keep the $(\text{num}, \text{den})$ pair separate; \texttt{MaynardVerify.v}
operates on those twins to avoid \texttt{rat}'s
canonical \texttt{gcd}-normalisation cost during \vmcompute.
\end{remark}

\section{Kernel cross-check: matrices match the spec}

\begin{theorem}[$M_1$ entries match Maynard's closed form]\label{thm:M1_match}
\rocq{PrimeGapS1.MaynardVerify.Def.all_match_M1Z_true}
\rocqok
\uses{def:M1_entry}
For every $(i, j) \in \{0, \ldots, 41\}^2$,
\[
\texttt{m1\_num}(i, j) \cdot \DMone
   \;=\; M_1^{\mathrm{int}}[i][j] \cdot \texttt{m1\_den}(i, j)\quad\text{in }\Z,
\]
where $M_1^{\mathrm{int}}$ is the FLINT-shipped integer matrix and
$(\texttt{m1\_num}, \texttt{m1\_den})$ are the closed-form
$(\text{numerator}, \text{denominator})$ from
Definition~\ref{def:M1_entry}.
\end{theorem}
\begin{proof}\rocqok
The $1764$ Boolean cross-multiplication checks
\texttt{M1\_entry\_matchZ}~$i\,j$ are aggregated into
\texttt{all\_match\_M1Z}~$=$~\texttt{forallb}~over $42 \times 42$, and
the Lemma is closed by a single \vmcompute{}. The check costs
${\sim}90$~s wall-clock. \texttt{Print Assumptions} reports only the
standard \Uintsf{} primitives.
\end{proof}

\begin{theorem}[$M_2$ entries match Maynard's closed form]\label{thm:M2_match}
\rocq{PrimeGapS1.MaynardVerify.all_match_M2Z_true}
\rocqok
\uses{def:M2_entry}
The analogous identity holds for $M_2$:
$\texttt{m2\_num}(i, j) \cdot \DMtwo
   = M_2^{\mathrm{int}}[i][j] \cdot \texttt{m2\_den}(i, j)$
for all $(i, j) \in \{0, \ldots, 41\}^2$.
\end{theorem}
\begin{proof}\rocqok
Same shape as Theorem~\ref{thm:M1_match}, but the entries are double
sums of up to 36 rational terms, so the per-row \vmcompute{} cost is
significant. The check is split into six 7-row chunks
(\texttt{MaynardVerify/M2\_0..5.v}) that \texttt{make -j} runs
concurrently, and the assembly stitches them via \texttt{seq\_split\_42}
plus \texttt{forallb\_app} — no per-entry recomputation in the
\Qed{} of \texttt{all\_match\_M2Z\_true} itself.
\texttt{Print Assumptions}: only the standard \Uintsf{} primitives.
\end{proof}

\begin{remark}
Theorems~\ref{thm:M1_match} and~\ref{thm:M2_match} together close the
trust gap on the input data: every entry of the $42 \times 42$ integer
matrices that the rest of the proof consumes is kernel-verified to
agree with Maynard's closed-form specification (Lemma~8.2). The FLINT
generator is therefore not part of the trust base for the input data
either.
\end{remark}

\section{Paper-form $\leftrightarrow$ computational-form spec bridge}

The cross-checks above operate on the $\Z$-level twins
(\texttt{m1\_num\_den}, \texttt{m2\_num\_den}) because \vmcompute{} on
$\rat$-level matrices stalls on canonical-structure elaboration
(see~\S\ref{ch:hb}). The rat-level spec
(\texttt{M1\_entry}, \texttt{M2\_entry}) is the
documentation-shaped form a reviewer reads side by side with
arXiv:1311.4600~\S8. The bridge below certifies in the kernel that the
two forms encode the same closed forms: a single \Qed{} per matrix,
with no axioms.

\begin{definition}[Reading a $\Z$-pair as a rational]\label{def:qfrac}
\rocq{PrimeGapS1.MaynardSpecBridge.qfrac}
\rocqok
For $(n, d) \in \Z \times \Z$, set
$\texttt{qfrac}(n, d) := \frac{n}{d} \in \rat$
(MathComp's int--to--rat morphism applied component-wise, then divided).
This reads a $(\text{numerator}, \text{denominator})$ pair as the
rational it represents.
\end{definition}

\begin{theorem}[$\Mone$ paper-form spec equals $\Z$-level spec]\label{thm:M1_spec_rat_eq}
\rocq{PrimeGapS1.MaynardSpecBridge.M1_spec_rat_eq}
\rocqok
\uses{def:M1_entry, def:qfrac}
For every $(i, j) \in \{0, \ldots, 41\}^2$,
\[
  \texttt{M1\_spec\_ij}(i, j) \;=\; \texttt{qfrac}\bigl(\texttt{m1\_num\_den\_at}(i, j)\bigr).
\]
\end{theorem}
\begin{proof}\rocqok
Layered structural induction in
\texttt{theories/S1/MaynardSpecBridge.v}: \texttt{factZ\_to\_rat},
\texttt{dblratZ\_to\_rat}, \texttt{binZ\_to\_rat},
\texttt{compositionsZ\_eq\_compositions}, \texttt{G2Z\_to\_rat}, and
\texttt{m1\_num\_den\_to\_rat} compose into a single rewrite chain.
\texttt{Print Assumptions} reports \emph{Closed under the global
context}: no axioms, not even \Uintsf\ primitives.
\end{proof}

\begin{theorem}[$\Mtwo$ paper-form spec equals $\Z$-level spec]\label{thm:M2_spec_rat_eq}
\rocq{PrimeGapS1.MaynardSpecBridge.M2_spec_rat_eq}
\rocqok
\uses{def:M2_entry, def:qfrac}
For every $(i, j) \in \{0, \ldots, 41\}^2$,
\[
  \texttt{M2\_spec\_ij}(i, j) \;=\; \texttt{qfrac}\bigl(\texttt{m2\_num\_den\_at}(i, j)\bigr).
\]
\end{theorem}
\begin{proof}\rocqok
Same layered induction, with two extra steps for the running
$\texttt{qplus}$ accumulator: \texttt{qfrac\_qmul} and
\texttt{qfrac\_qplus} bridge \texttt{qmul} / \texttt{qplus} on
$\Z \times \Z$ pairs to $*$ / $+$ on $\rat$, and a generic
$\texttt{fold\_left qplus} \to \sum$ lemma collects the double sum
over $(c'_1, c'_2)$. \texttt{Print Assumptions}: \emph{Closed under
the global context}.
\end{proof}

\begin{remark}
Theorems~\ref{thm:M1_spec_rat_eq} and~\ref{thm:M2_spec_rat_eq} live
in \texttt{MaynardSpecBridge.v}, a leaf in the dependency DAG:
\texttt{Cert.v} does not import them, and the headline theorem does
not depend on them. Their job is purely to certify that the
documentation-shaped $\rat$-level definitions
(Definitions~\ref{def:M1_entry} and~\ref{def:M2_entry}, written to
match Maynard's paper character for character) and the
$\vmcompute$-shaped $\Z$-level definitions
(\texttt{m1\_num\_den}, \texttt{m2\_num\_den}, consumed by
Theorems~\ref{thm:M1_match} and~\ref{thm:M2_match}) compute the same
rational. Combined with those theorems, this gives a two-step
kernel-verified path from each shipped integer matrix entry back to
Maynard's exact closed form.
\end{remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{The characteristic polynomial route}\label{ch:charpoly}

This chapter develops the in-kernel characteristic polynomial
machinery.  We use a hand-rolled Faddeev--LeVerrier (FL) recurrence
on \texttt{list (list Z)} matrices and prove that it computes the
same polynomial as MathComp's \texttt{char\_poly} on the
corresponding $\Q$-matrix.  The material lives in
\texttt{IntPoly.v}, \texttt{IntMat.v}, and \texttt{CharPoly.v}.

\section{Integer polynomial and matrix libraries}

\begin{definition}[Integer polynomial type]\label{def:pol}
\rocq{PrimeGapS1.IntPoly.pol}
\rocqok
$\texttt{pol} := \texttt{list }\Z$, low-to-high coefficient order. The
operations \texttt{padd}, \texttt{psub}, \texttt{pneg}, \texttt{pscale},
\texttt{pmul}, \texttt{pderiv}, \texttt{prem} (pseudo-remainder),
\texttt{plead}, \texttt{pdeg}, \texttt{pnorm} (strip trailing zeros),
\texttt{peval} and \texttt{peval\_at\_rat} are all
\vmcompute-reducible and proved correct against ordinary integer
arithmetic.
\end{definition}

\begin{definition}[Integer matrix type]\label{def:mat}
\rocq{PrimeGapS1.IntMat.mat}
\rocqok
$\texttt{mat} := \texttt{list (list }\Z\texttt{)}$, row-major, no
well-formedness invariant. The operations \texttt{mzero},
\texttt{meye}, \texttt{madd}, \texttt{mscale}, \texttt{mmul},
\texttt{mtrans}, \texttt{mtrace}, \texttt{mat\_get} are pure list
recursions; nothing imports MathComp at this layer. The reason is that
MathComp's $'M[\Q]_{42}$ does not reduce under \vmcompute{} at the
sizes we need, but nested \texttt{list (list Z)} does
(${\sim}0.14$~s for a $42 \times 42$ multiplication).
\end{definition}

\section{Faddeev--LeVerrier on \texorpdfstring{$\Z$}{Z}-matrices}

\begin{definition}[FL recurrence]\label{def:fl_loop}
\rocq{PrimeGapS1.CharPoly.fl_loop, PrimeGapS1.CharPoly.char_poly_int}
\rocqok
\uses{def:mat, def:pol}
\texttt{fl\_loop} runs the classical Faddeev--LeVerrier recurrence
$M_0 := 0$, $c_0 := 1$, $M_k := A \cdot M_{k-1} + c_{k-1} I$,
$c_k := -\trace(A \cdot M_k)/k$, accumulating the coefficient list.
$\texttt{char\_poly\_int}(A) := \texttt{rev}(c_0, c_1, \ldots, c_n) \mathrel{++} [1]$
gives the monic char poly with leading coefficient~$1$.
\end{definition}

\begin{lemma}[FL agrees with $\Q$-FL]\label{lem:fl_loop_rat_is_charpoly}
\rocq{PrimeGapS1.CharPoly.fl_loop_rat_is_char_poly_L2,
      PrimeGapS1.CharPoly.fl_invariant_L2,
      PrimeGapS1.CharPoly.fl_divisibility_L2}
\rocqok
\uses{def:fl_loop}
After viewing $A : \texttt{mat}$ as the $\Q$-matrix
$\texttt{mat\_int\_to\_rat}(A, 1, n) : 'M[\Q]_n$, the integer FL
recurrence reproduces the rational FL recurrence step by step. The
auxiliary invariants \texttt{fl\_invariant\_L2} (the
relationship between $M_k$, $c_k$ and the rational $M_k^{\Q}$) and
\texttt{fl\_divisibility\_L2} (the classical Newton-identity fact that
$\trace(A \cdot M_k)$ is divisible by~$k$ at every step) are both
proved by induction on the FL step count.
\end{lemma}
\begin{proof}\rocqok
Long induction. The hard part is \texttt{fl\_divisibility\_L2}: the
$k$-th Newton power-sum identity
$p_k = \sum_{i=1}^{k-1}(-1)^{i-1} e_i p_{k-i} + (-1)^{k-1} k e_k$
is the algebraic content. We use the explicit recurrence form rather
than the global identity, which lets the induction track each FL step
in lockstep.
\end{proof}

\begin{theorem}[FL correctness over $\Q$]\label{thm:char_poly_int_correct}
\rocq{PrimeGapS1.CharPoly.char_poly_int_correct}
\rocqok
\uses{lem:fl_loop_rat_is_charpoly}
For a square integer matrix $M$ of dimension $n$,
\[
\texttt{pol\_to\_polyrat}(\texttt{char\_poly\_int}(M))
   \;=\; \texttt{char\_poly}(\texttt{mat\_int\_to\_rat}(M, 1, n))
   \quad\text{in } \Q[X].
\]
\end{theorem}
\begin{proof}\rocqok
Combine Lemma~\ref{lem:fl_loop_rat_is_charpoly} with the bridging
definitions \texttt{Z\_to\_int}, \texttt{mat\_int\_to\_rat},
\texttt{pol\_to\_polyrat}: each FL step on the rational side equals
the standard MathComp \texttt{char\_poly} computation by
\texttt{Newton-identities-charpoly}. The induction is over the matrix
dimension; at the inductive step we use that the leading coefficient
of \texttt{char\_poly} is~$1$ and that the next-to-leading coefficient
is $-\trace M$, both already in MathComp.
\end{proof}

\section{The determinant-pencil identity}

The bridge from the two integer determinants
$\det(M_1^{\mathrm{int}})$ and $\det(\texttt{pencil\_int\_clean})$
to a sign of $\charpolysym_{\Arat}$ at $4/105$ is the
\emph{pencil identity}, a one-line mathcomp fact.

\begin{theorem}[Pencil determinant identity]\label{thm:det_pencil}
\rocq{PrimeGapS1.DetPencil.det_pencil}
\rocqok
For any commutative ring with unit $R$, dimension $n$, matrices
$M_1, M_2 : 'M[R]_n$ with $M_1 \in \mathrm{unitmx}$, and scalar
$\lambda$:
\[
\det\bigl(\lambda\,M_1 - M_2\bigr)
   \;=\; \det(M_1)\,\cdot\,\charpolysym\bigl(M_1^{-1}\,M_2\bigr)(\lambda).
\]
\end{theorem}
\begin{proof}\rocqok
One-line proof on top of \texttt{char\_poly\_horner\_eval}: substitute
$M_2 = (M_1)(M_1^{-1}M_2)$ in $\lambda M_1 - M_2$, factor $M_1$ out
of the determinant, and recognise the residual as
$\charpolysym(M_1^{-1}M_2)$ evaluated at $\lambda$.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{CRT cross-validation}\label{ch:crt}

The most computationally heavy step in the entire proof is showing
that the two shipped integer determinants
$\texttt{det\_M1\_int\_value}$ and $\texttt{D\_pencil\_int\_value}$
are the actual constant coefficients of
$\texttt{char\_poly\_int}(M_1^{\mathrm{int}})$ and
$\texttt{char\_poly\_int}(\texttt{pencil\_int\_clean})$
respectively.  Both integers are large ($2044$ and $2613$ bits) and
direct \vmcompute{} on the LHS would take hours and produce
unwieldy proof terms.  Instead we use a Chinese Remainder Theorem
(CRT) lift: verify the equality modulo $710$ Uint63 primes, then
close $|a - b| = 0$ from $(\prod p_i) \mid (a - b)$ via a verified
Hadamard-style coefficient bound.  This chapter follows files
\texttt{ModularArith.v}, \texttt{CRTBridge.v},
\texttt{CharPolyAgree/Def.v}, \texttt{CRTCheck.v},
\texttt{Fermat.v}, \texttt{PrimeCheck.v}, and \texttt{CRTLift.v}.

\section{Modular arithmetic in Uint63}

\begin{definition}[Modular FL char-poly]\label{def:char_poly_mod}
\rocq{PrimeGapS1.ModularArith.char_poly_mod}
\rocqok
\uses{def:fl_loop}
\texttt{char\_poly\_mod}~$p\,M$ runs the same FL recurrence as
$\texttt{char\_poly\_int}$ but modulo a single $\Uintsf$ prime $p$,
using only native 63-bit operations: \texttt{addmod63},
\texttt{mulmod63}, \texttt{negmod63}, plus modular inverse
\texttt{inv\_mod63}~$p\,a := a^{p - 2} \bmod p$ (Fermat's little
theorem). For $p < 2^{31}$, all intermediate products $a \cdot b$ fit
in $2^{62} < 2^{63}$, so no overflow occurs and \vmcompute{}
processes a single $42 \times 42$ FL chain in ${\sim}0.5$~s.
\end{definition}

\begin{remark}[Why \texttt{ModularArith.v} exists as one file]
\label{rem:modular_arith_factor}
The original development duplicated all modular operations across
\texttt{CRTBridge.v} and the per-prime check files.  When
\texttt{CRTPencilCheck.v} tried to rewrite a fact mentioning
\texttt{char\_poly\_mod}\,$p\,M_1^{\mathrm{int}}$ from one file by
an equation mentioning the same expression but defined in another
file, the kernel saw two distinct head constants and tried to
delta-unfold both, which is exactly the FL recurrence on a
concrete $42 \times 42$ matrix.  Empirically that hung the kernel
for $> 25$ minutes per rewrite.  Factoring every modular operation
into one file pins each constant to a single kernel name, so
unification of the same head symbol is by reflexivity.
\end{remark}

\section{Per-prime correctness of modular FL}

\begin{theorem}[Soundness of \texttt{char\_poly\_mod}]\label{thm:char_poly_mod_sound}
\rocq{PrimeGapS1.CRTBridge.char_poly_mod_sound}
\rocqok
\uses{def:char_poly_mod, def:fl_loop}
For any $\Uintsf$ prime $p$ with $1 < p < 2^{31}$ and $n + 1 < p$
(where $n$ is the matrix dimension), and any \texttt{fl\_all\_divisible}
matrix~$M$,
\[
\texttt{List.map}\,(\texttt{Z\_to\_mod63}\,p)\,(\texttt{char\_poly\_int}\,M)
   \;=\; \texttt{char\_poly\_mod}\,p\,M.
\]
The hypothesis $n + 1 < p$ ensures every FL division by $k \in \{1, \ldots, n\}$
is well-defined in $\Fp$.
\end{theorem}
\begin{proof}\rocqok
Thread soundness lemmas for each modular operation through the FL
recurrence. The base step is $\texttt{Z\_to\_mod63\_spec}$, which
proves
$\texttt{to\_Z}(\texttt{Z\_to\_mod63}\,p\,z) = z \bmod \texttt{to\_Z}\,p$
under \texttt{valid\_prime}~$p$ (precisely the
$2^{31}$-bound that guarantees no $\Uintsf$ multiplication overflow).
The inductive step combines $\texttt{addmod63}$/$\texttt{mulmod63}$
soundness with the FL invariants from
Lemma~\ref{lem:fl_loop_rat_is_charpoly}, plus Fermat's little theorem
(below) to discharge the modular division.
\end{proof}

\begin{lemma}[Fermat's little theorem, modular inverse form]
\label{lem:fermat_mod}
\rocq{PrimeGapS1.Fermat.fermat_mod, PrimeGapS1.Fermat.fermat_Z}
\rocqok
For prime $p$ and $1 \le k < p$,
$k \cdot k^{p - 2} \equiv 1 \pmod{p}$.
The $\Z$-level twin \texttt{fermat\_Z} bridges to
\texttt{Znumtheory.prime}; together with
\texttt{Zprime\_to\_ssrprime} this provides the interface that the
710-prime CRT lift consumes.
\end{lemma}
\begin{proof}\rocqok
Via MathComp's \texttt{expf\_card} on the multiplicative group $\Fp^\ast$.
\end{proof}

\section{The 710 primes}

\begin{definition}[CRT prime list]\label{def:crt_primes}
\rocq{PrimeGapS1.CharPolyAgree.Def.crt_primes_local,
      PrimeGapS1.CharPolyAgree.Def.crt_primes_extra}
\rocqok
A concatenated list \texttt{crt\_primes\_all} of $710$ distinct
$\Uintsf$ primes, each in $[2^{30}, 2^{31})$. The product
$\prod_{i=1}^{710} p_i$ exceeds $2^{21300}$.
\end{definition}

\begin{lemma}[All 710 candidates are prime]\label{lem:crt_primes_all_prime}
\rocq{PrimeGapS1.CRTLift.crt_primes_710_all_prime,
      PrimeGapS1.CRTLift.crt_primes_710_NoDup}
\rocqok
\uses{def:crt_primes}
Every entry of \texttt{crt\_primes\_all} is a prime number, and the
list contains no duplicates.
\end{lemma}
\begin{proof}\rocqok
Primality: \texttt{check\_prime\_Z}, a pure-$\Z$ trial-division checker
proved sound for \texttt{Znumtheory.prime} in \texttt{PrimeCheck.v}, is
applied to each entry by \texttt{forallb} and reduced by \vmcompute.
Each check costs ${\sim}0.6$~s (trial division up to
$\lfloor\sqrt{p}\rfloor \approx 2^{15}$), so all $710$ checks run in
${\sim}7$~min. The $\texttt{NoDup}$ check is a quadratic
\texttt{nodup\_Z} also discharged by \vmcompute. The latter is bridged
to MathComp's \texttt{ssrnat.prime} via
\texttt{Zprime\_to\_ssrprime}.
\end{proof}

\section{The CRT lift}

\begin{lemma}[Small multiple is zero]\label{lem:small_multiple_zero}
\rocq{PrimeGapS1.CRTCheck.small_multiple_zero}
\rocqok
If $P \mid c$, $0 < P$, and $2|c| < P$ (all in $\Z$), then $c = 0$.
\end{lemma}
\begin{proof}\rocqok
A divisor of $c$ has absolute value $\le |c|$ unless $c = 0$. If
$c \ne 0$ then $P \le |c| < P/2$, contradiction. \texttt{lia} closes
the integer arithmetic.
\end{proof}

\begin{lemma}[Distinct primes divide the product]\label{lem:all_primes_divide}
\rocq{PrimeGapS1.CRTCheck.all_primes_divide_product}
\rocqok
For a no-duplicate list \texttt{ps} of primes, if $p \mid c$ for every
$p \in \texttt{ps}$, then $\prod\texttt{ps} \mid c$.
\end{lemma}
\begin{proof}\rocqok
Induction on \texttt{ps}. The inductive step uses the elementary fact
that distinct primes are coprime, hence $p \mid c$ and $q \mid c$
imply $pq \mid c$ when $\gcd(p, q) = 1$. The \texttt{NoDup} hypothesis
is precisely what guarantees pairwise coprimality at each step.
\end{proof}

\begin{theorem}[CRT lift: integer determinants match shipped literals]
\label{thm:det_int_eq}
\rocq{PrimeGapS1.CRTPencilCheck.det_M1_int_eq,
      PrimeGapS1.CRTPencilCheck.D_pencil_int_eq}
\rocqok
\uses{def:char_poly_mod, thm:char_poly_mod_sound, def:crt_primes,
      lem:crt_primes_all_prime, lem:small_multiple_zero,
      lem:all_primes_divide}
$\texttt{det\_M1\_int} = \texttt{det\_M1\_int\_value}$ and
$\texttt{D\_pencil\_int} = \texttt{D\_pencil\_int\_value}$, both
in $\Z$, where
$\texttt{det\_M1\_int}$ /
$\texttt{D\_pencil\_int}$ are the sealed
$\sigT/\mathrm{proj1\_sig}$ wrappers around
$\texttt{List.nth 0 (char\_poly\_int}\,M_1^{\mathrm{int}})\,0$
and $\texttt{List.nth 0 (char\_poly\_int}\,\texttt{pencil\_int\_clean})\,0$
respectively.
\end{theorem}
\begin{proof}\rocqok
The two proofs share the same template; we describe the $M_1$ case.
We reduce $a - b = 0$ (with
$a := \texttt{det\_M1\_int}$, $b := \texttt{det\_M1\_int\_value}$)
to Lemma~\ref{lem:small_multiple_zero}.

For each prime $p$ in \texttt{crt\_primes\_all},
Theorem~\ref{thm:char_poly_mod_sound} gives
$\texttt{List.map}\,(\texttt{Z\_to\_mod63}\,p)\,
   (\texttt{char\_poly\_int}\,M_1^{\mathrm{int}})
   = \texttt{char\_poly\_mod}\,p\,M_1^{\mathrm{int}}$.
A single-prime \vmcompute{} check inside
\texttt{check\_M1\_det\_710\_true} (a Boolean check that all $710$
per-prime agreements hold simultaneously, closed by \vmcompute{}
$+$ \texttt{reflexivity}) gives
$\texttt{List.nth 0 (char\_poly\_mod}\,p\,M_1^{\mathrm{int}})\,0
   = \texttt{Z\_to\_mod63}\,p\,\texttt{det\_M1\_int\_value}$.
Combining the two, $p \mid (a - b)$.  By
Lemma~\ref{lem:all_primes_divide}, $\prod p \mid (a - b)$.

The bound $|a|, |b| \le B$ is established by a Hadamard-style
chain (\texttt{CRTPencilHadamardGeneric.v}) applied to
$M_1^{\mathrm{int}}$, then specialised in
\texttt{CRTPencilM1Bound.v}; the closed-form
$\texttt{fl\_coeff\_bound}\,42\,(\texttt{max\_abs\_entry}\,M_1^{\mathrm{int}})$
matches the shipped literal $\texttt{fl\_coeff\_bound\_M1\_value}$
via the \vmcompute{}-Qed
$\texttt{fl\_coeff\_bound\_M1\_eq}$.
The combined bound $2B + 2|\texttt{det\_M1\_int\_value}|
< \prod p_i$ is verified by \vmcompute{} in
$\texttt{crt\_bound\_M1\_sufficient\_literal}$.
Lemma~\ref{lem:small_multiple_zero} then closes $a = b$.

The pencil case (\texttt{D\_pencil\_int\_eq}) is structurally
identical, with
\texttt{CRTPencilChecksProof.check\_pencil\_det\_710\_true},
\texttt{CRTPencilPencilBound.crt\_bound\_pencil\_sufficient\_literal},
and $\texttt{pencil\_int\_clean}$ in place of their $M_1$
counterparts.  The clean-pencil refactor was what made the same
$710$-prime product suffice for both determinants.
\end{proof}

\begin{remark}[\texttt{Strategy opaque} as a kernel-engineering tool]
\label{rem:strategy_opaque}
A naive proof of the per-prime fact
$\texttt{List.nth 0 (char\_poly\_mod}\,p\,M_1^{\mathrm{int}})\,0
   = \texttt{Z\_to\_mod63}\,p\,\texttt{det\_M1\_int\_value}$
hangs the \Qed{} at conversion time: the kernel's ordinary reducer
forces WHNF on $\texttt{char\_poly\_mod}\,p\,M_1^{\mathrm{int}}$,
which triggers the full FL recurrence on a concrete $42 \times 42$
matrix.  The fix is to wrap the \Qed{} in
\texttt{Strategy opaque [char\_poly\_mod}\,$M_1^{\mathrm{int}}$\,\texttt{\ldots]},
forcing the kernel to compare head symbols structurally, not unfold
the values.  This is mathematically a no-op but operationally
indispensable.  The shipped determinants \texttt{det\_M1\_int} and
\texttt{D\_pencil\_int} are additionally sealed via
\texttt{sigT}/\texttt{proj1\_sig} in
\texttt{theories/S1/CertPencilDef.v} so that the conversion-time
opacity propagates to all downstream uses.
\end{remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{The pencil-determinant assembly}\label{ch:pencil}

This chapter assembles the eigenvalue bound from the pencil
identity (Theorem~\ref{thm:det_pencil}), the two integer
determinant lifts
(\texttt{CRTPencilCheck.det\_M1\_int\_eq} and
\texttt{D\_pencil\_int\_eq}), the per-cell clean-pencil grid
identity
(\texttt{PencilCleanGrid.all\_pencil\_clean\_match\_true}), and the
intermediate value theorem.  The material lives in
\texttt{DetPencil.v}, \texttt{CertPencilDef.v},
\texttt{AbstractPencilHelper.v}, \texttt{PencilCleanGrid.v},
\texttt{CRTPencilCheck.v}, \texttt{CRTPencilChecksProof.v},
\texttt{CRTPencilHadamardGeneric.v}, \texttt{CRTPencilM1Bound.v},
\texttt{CRTPencilPencilBound.v}, and \texttt{CertPencil.v}.

\section{Signs of the two integer determinants}

\begin{lemma}[Sign of $\det(M_1^{\mathrm{int}})$]\label{lem:det_M1_int_pos}
\rocq{PrimeGapS1.CertPencil.det_M1_int_pos,
      PrimeGapS1.CRTPencilCheck.det_M1_int_eq}
\rocqok
The shipped integer determinant
\texttt{det\_M1\_int\_value} (in
\texttt{theories/S1/Witness\_PencilDet.v}) equals the constant
coefficient of $\texttt{char\_poly\_int}(M_1^{\mathrm{int}})$, and
it is strictly positive.
\end{lemma}
\begin{proof}\rocqok
The equality follows from the $710$-prime CRT lift of
Chapter~\ref{ch:crt} applied to the shipped literal:
\texttt{check\_M1\_det\_710\_true} checks
$\texttt{List.nth 0 (char\_poly\_mod}\,p\,M_1^{\mathrm{int}})\,0
  = \texttt{Z\_to\_mod63}\,p\,\texttt{det\_M1\_int\_value}$
modulo each of the $710$ primes via \vmcompute, and a closed-form
Hadamard bound
(\texttt{CRTPencilM1Bound.crt\_bound\_M1\_sufficient\_literal})
ensures $2 \cdot \texttt{fl\_coeff\_bound}\,42\,(\texttt{max\_abs\_entry}\,M_1^{\mathrm{int}})
+ 2 \cdot |\texttt{det\_M1\_int\_value}|
<\texttt{crt\_product\_710}$.  The determinant literal is
$2044$ bits and the CRT product is ${\sim}21300$ bits, so the
inequality holds by \vmcompute.  Positivity of the literal is a
single-\vmcompute{} Qed
(\texttt{det\_M1\_int\_value\_pos}).
\end{proof}

\begin{lemma}[Sign of $\det(\texttt{pencil\_int\_clean})$]\label{lem:D_pencil_neg}
\rocq{PrimeGapS1.CertPencil.D_pencil_int_neg,
      PrimeGapS1.CRTPencilCheck.D_pencil_int_eq}
\rocqok
The shipped integer determinant
\texttt{D\_pencil\_int\_value} (in
\texttt{theories/S1/Witness\_PencilClean.v}) equals the constant
coefficient of
$\texttt{char\_poly\_int}(\texttt{pencil\_int\_clean})$, and it is
strictly negative.
\end{lemma}
\begin{proof}\rocqok
Same CRT pattern as Lemma~\ref{lem:det_M1_int_pos}, with the
Hadamard bound discharged by
\texttt{CRTPencilPencilBound.crt\_bound\_pencil\_sufficient\_literal}.
The pencil determinant literal is $2613$ bits, comfortably under
the same $710$-prime CRT product.  Negativity is a single
\vmcompute{} Qed (\texttt{D\_pencil\_int\_value\_neg}).
\end{proof}

\section{Per-cell clean-pencil identity}

\begin{lemma}[Per-cell clean-pencil match]\label{lem:pencil_clean_match}
\rocq{PrimeGapS1.PencilCleanGrid.all_pencil_clean_match_true,
      PrimeGapS1.PencilCleanGrid.pencil_clean_match_Z}
\rocqok
For every $(i, j) \in \{0, \ldots, 41\}^2$,
\[
D_{M_1}\,D_{M_2}\,\texttt{pencil\_int\_clean}[i, j]
   \;=\; D_{\mathrm{pencil}}^{\mathrm{clean}}\,\cdot\,
   \bigl(4\,D_{M_2}\,M_1^{\mathrm{int}}[i, j]
         - 105\,D_{M_1}\,M_2^{\mathrm{int}}[i, j]\bigr)
\quad\text{in } \Z.
\]
\end{lemma}
\begin{proof}\rocqok
Single 1764-cell \vmcompute.  The literal
$D_{\mathrm{pencil}}^{\mathrm{clean}}$ is the LCM of the
denominators of $4 M_1 - 105 M_2$ (689~bits, roughly half of
$D_{M_1} \cdot D_{M_2} = 1368$~bits), so the resulting integer
matrix \texttt{pencil\_int\_clean} stays comfortably under the
710-prime CRT product.
\end{proof}

\section{The pencil sign of \texorpdfstring{$\charpolysym_{\Arat}(4/105)$}{chi A\_rat at 4/105}}

\begin{theorem}[Pencil sign of $\charpolysym_{\Arat}(4/105)$]\label{thm:charpoly_neg}
\rocq{PrimeGapS1.CertPencil.abstract_charpoly_neg,
      PrimeGapS1.CertPencil.charpoly_neg_at_threshold_rat}
\rocqok
\uses{thm:det_pencil, lem:det_M1_int_pos, lem:D_pencil_neg,
      lem:pencil_clean_match}
The rational evaluation $(\charpolysym_{\Arat}).\![4/105] < 0$.
\end{theorem}
\begin{proof}\rocqok
By the determinant-pencil identity (Theorem~\ref{thm:det_pencil})
specialised at $M_1 := \Mone^{\mathrm{rat}}$,
$M_2 := \Mtwo^{\mathrm{rat}}$, $\lambda := 4/105$:
\[
\det\!\bigl((4/105)\,M_1 - M_2\bigr)
  = \det(M_1) \cdot (\charpolysym_{\Arat}).\![4/105].
\]
Lemma~\ref{lem:pencil_clean_match} lifts to the rat-level matrix
equality
$D_{M_1} D_{M_2} (4 M_1^{\mathrm{rat}} - 105 M_2^{\mathrm{rat}})
  = D_{\mathrm{pencil}}^{\mathrm{clean}}\,\cdot\,\texttt{pencil\_int\_clean}$
(via \texttt{AbstractPencilHelper.pencil\_matrix\_bridge} at
generic dimension), so
$\det\!\bigl((4/105) M_1 - M_2\bigr)$ has the same sign as
$\det(\texttt{pencil\_int\_clean})$, which is negative by
Lemma~\ref{lem:D_pencil_neg}.  Lemma~\ref{lem:det_M1_int_pos}
gives $\det(M_1) > 0$, so the right-hand side of the pencil
identity has the same sign as the left, and therefore
$(\charpolysym_{\Arat}).\![4/105] < 0$.
\end{proof}

\section{IVT: a root exists above \texorpdfstring{$4/105$}{4/105}}

\begin{theorem}[Root in realalg above $4/105$]\label{thm:maynard_root}
\rocq{PrimeGapS1.CertPencil.maynard_root_above_threshold}
\rocqok
\uses{thm:charpoly_neg}
There exists $\lambda \in \realalg$ such that
$(\texttt{map\_poly}\,\texttt{ratr}\,\charpolysym_{\Arat}).[\lambda] = 0$
and $\texttt{ratr}(4/105) < \lambda$.
\end{theorem}
\begin{proof}\rocqok
MathComp's \texttt{poly\_ivtoo} on a real-closed field.  Two
endpoint sign facts on
$\widehat{P} := \texttt{map\_poly}\,\texttt{ratr}\,\charpolysym_{\Arat}$:
\begin{itemize}
\item Lower endpoint $\texttt{ratr}(4/105)$: by
  Theorem~\ref{thm:charpoly_neg} together with the $\texttt{ratr}$
  morphism pushed through \texttt{horner}.
\item Upper endpoint $\texttt{cauchy\_bound}\,\widehat{P}$: the
  leading coefficient of $\charpolysym_{\Arat}$ is $1 > 0$ (so the
  leading coefficient of $\widehat{P}$ is also $1 > 0$); by the
  standard property of \texttt{cauchy\_bound}, $\widehat{P}$
  evaluated above its Cauchy bound has the sign of its leading
  coefficient, hence is positive.
\end{itemize}
\texttt{poly\_ivtoo} yields $x \in [\texttt{ratr}(4/105),
\texttt{cauchy\_bound}\,\widehat{P}]$ with $\widehat{P}(x) = 0$;
strictness at the lower endpoint follows because $\widehat{P}$ is
\emph{strictly} negative there.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Assembly: from root to eigenvalue}\label{ch:assembly}

This chapter glues the IVT root of Chapter~\ref{ch:pencil}, the
CRT lift of Chapter~\ref{ch:crt}, and the FL correctness of
Chapter~\ref{ch:charpoly} into the headline theorem.  The two-line
\Qed{} of \texttt{maynard\_eigenvalue\_S1\_pencil} in
\texttt{CertPencil.v} is what holds the whole proof tree together.

\section{\texorpdfstring{$\Arat$ is a well-defined $\Q$-matrix}{A\_rat is a well-defined Q-matrix}}

\begin{definition}[The rational pencil matrix]\label{def:A_rat}
\rocq{PrimeGapS1.CertL2.A_rat}
\rocqok
\uses{def:M1_entry, def:M2_entry}
\[
\Arat \;:=\; \mathrm{invmx}(\texttt{mat\_int\_to\_rat}(\Mone^{\mathrm{int}}, \DMone, 42))
            \;\cdot\; \texttt{mat\_int\_to\_rat}(\Mtwo^{\mathrm{int}}, \DMtwo, 42)
\quad\in\; 'M[\Q]_{42}.
\]
This is the matrix whose largest eigenvalue we want to lower-bound by
$4/105$.
\end{definition}

\begin{lemma}[$\Mone$ is invertible over $\Q$]\label{lem:M1_unit}
\rocq{PrimeGapS1.CertL2.M1_1_unit}
\rocqok
$\texttt{mat\_int\_to\_rat}(\Mone^{\mathrm{int}}, 1, 42) \in \mathrm{unitmx}$.
\end{lemma}
\begin{proof}\rocqok
A modular determinant check at one of the CRT primes $p$ shows
$\det \Mone^{\mathrm{int}} \not\equiv 0 \pmod p$
(via $\texttt{char\_poly\_mod}$ — the constant term of the modular
char poly is $\pm \det$, so a nonzero constant term modulo any prime
witnesses $\det \ne 0$ over $\Z$, hence over $\Q$). MathComp's
\texttt{unitmxE} closes the goal from $\det \ne 0$.
\end{proof}

\section{The headline theorem}

\begin{lemma}[Char-poly root $\Rightarrow$ eigenvalue]
\label{lem:eigenvalue_of_root}
\rocq{PrimeGapS1.CertPencil.maynard_eigenvalue_S1_pencil}
\rocqok
For any $\lambda \in \realalg$, if
$\texttt{root}\,(\texttt{map\_poly}\,\texttt{ratr}\,\charpolysym(\Arat))\,\lambda$
then $\texttt{eigenvalue}\,(\texttt{map\_mx}\,\texttt{ratr}\,\Arat)\,\lambda$.
\end{lemma}
\begin{proof}\rocqok
A two-line script. \texttt{map\_char\_poly} (a MathComp lemma) says
$\charpolysym(\texttt{map\_mx}\,f\,M) = \texttt{map\_poly}\,f\,(\charpolysym\,M)$
for any ring morphism $f$; \texttt{eigenvalue\_root\_char} (also
MathComp) says an eigenvalue is exactly a root of the char poly.
Composing the two gives the implication.
\end{proof}

\begin{theorem}[Spectral headline: an eigenvalue exceeds $4/105$]
\label{thm:headline}
\rocq{PrimeGapS1.CertPencil.maynard_eigenvalue_S1_pencil}
\rocqok
\uses{thm:maynard_root, lem:eigenvalue_of_root}
There exists $\lambda \in \realalg$ such that
\[
\texttt{eigenvalue}\,(\texttt{map\_mx}\,\texttt{ratr}\,\Arat)\,\lambda
\quad\text{and}\quad
\texttt{ratr}(4/105) < \lambda.
\]
\end{theorem}
\begin{proof}\rocqok
Two-line tactic.  Theorem~\ref{thm:maynard_root} produces
$\lambda \in \realalg$ that is a root of
$\texttt{map\_poly}\,\texttt{ratr}\,\charpolysym(\Arat)$ and
satisfies $\texttt{ratr}(4/105) < \lambda$.
Lemma~\ref{lem:eigenvalue_of_root} converts a char-poly root to an
eigenvalue of $\texttt{map\_mx}\,\texttt{ratr}\,\Arat$.  The strict
inequality is preserved verbatim.  Combining gives the existential.
\end{proof}

\begin{theorem}[End-to-end certification]\label{thm:M105}
\rocq{PrimeGapS1.CertPencil.maynard_M105_certified_pencil}
\rocqok
\uses{lem:M1_spec_eq_int, lem:M2_spec_eq_int, thm:headline}
The conjunction, in the main-branch shape ($M_{105} := 105\cdot\Arat$),
\[
\begin{aligned}
&\texttt{matches\_closed\_forms}(M_{105})\\
&\quad\wedge\;
\bigl(\exists \lambda \in \realalg.\;
   \texttt{eigenvalue}(\texttt{map\_mx}\,\texttt{ratr}\,M_{105})\,\lambda
   \;\wedge\;
   4 < \lambda\bigr),
\end{aligned}
\]
where $\texttt{matches\_closed\_forms}(M)$ abbreviates the three-fold
conjunction
\[
\begin{aligned}
&M = M_{105},\\
&\bigl(\forall i, j < 42.\;
  \texttt{M1\_spec\_ij}(i,j)
    = \texttt{Z2rat}(M_1^{\mathrm{int}}[i][j]) \,/\, \texttt{Z2rat}(\DMone)\bigr),\\
&\bigl(\forall i, j < 42.\;
  \texttt{M2\_spec\_ij}(i,j)
    = \texttt{Z2rat}(M_2^{\mathrm{int}}[i][j]) \,/\, \texttt{Z2rat}(\DMtwo)\bigr),
\end{aligned}
\]
and $\texttt{Z2rat}\,(z : \Z) : \rat := (\texttt{Z\_to\_int}\,z)\%\texttt{:\~{}R}$
is the thin $\Z \to \rat$ embedding. This is the same shape as the
\texttt{main} branch's \texttt{maynard\_M105\_certified}, stated over
$\realalg$ instead of \texttt{algC}.
\end{theorem}
\begin{proof}\rocqok
A two-way \texttt{split}. The $\texttt{matches\_closed\_forms}(M_{105})$
conjunct is itself a three-way split: the definitional
$M_{105} = 105\cdot\Arat$ (by \texttt{reflexivity}), and the two
composed per-matrix identities discharged by
Lemmas~\ref{lem:M1_spec_eq_int} and~\ref{lem:M2_spec_eq_int} (each one
identity per matrix, paper-form spec entry $=$ FLINT integer entry over
the common denominator). The eigenvalue existential is the spectral
Theorem~\ref{thm:headline} rescaled from $(\Arat, 4/105)$ to
$(M_{105}, 4)$ — multiply both the matrix and the threshold through by
$105$ — via \texttt{maynard\_eigenvalue\_S1\_pencil\_M105}. The two composed identities each fold
together a $\Z$-level cross-multiplication
(\texttt{all\_match\_M\{1,2\}Z\_true}) and a $\rat\leftrightarrow\Z$
bridge (\texttt{MaynardSpecBridge.M\{1,2\}\_spec\_rat\_eq}) via a thin
\texttt{qfrac\_eq\_div} helper. The Z-level booleans and the
$\rat\leftrightarrow\Z$ bridges remain available as standalone \Qed{}s
(Theorems~\ref{thm:M1_match}, \ref{thm:M2_match},
\ref{thm:M1_spec_rat_eq}, \ref{thm:M2_spec_rat_eq}) for auditors who
want to trace the chain step by step; they are not surfaced in the
headline statement. The combined \texttt{Print Assumptions} reports
only the standard Rocq \texttt{PrimInt63} kernel primitives (used by
\vmcompute{} on the $\Uintsf$ CRT computations). No project-specific
axioms.
\end{proof}

\begin{lemma}[$\Mone$ paper-form spec equals FLINT entry over $\DMone$]\label{lem:M1_spec_eq_int}
\rocq{PrimeGapS1.Cert.M1_spec_eq_int}
\rocqok
\uses{thm:M1_match, thm:M1_spec_rat_eq}
For every $(i, j) \in \{0, \ldots, 41\}^2$,
\[
  \texttt{M1\_spec\_ij}(i, j)
    \;=\; \texttt{Z2rat}\bigl(M_1^{\mathrm{int}}[i][j]\bigr) \,/\,
          \texttt{Z2rat}(\DMone) \quad\text{in } \rat.
\]
\end{lemma}
\begin{proof}\rocqok
Compose Theorem~\ref{thm:M1_spec_rat_eq}, which rewrites the
$\rat$-level spec into $\texttt{qfrac}(\texttt{m1\_num\_den\_at}(i,j))$,
with the $\Z$-level cross-multiplication from
Theorem~\ref{thm:M1_match} (extracted per-entry from the $42 \times 42$
\texttt{forallb}) via the helper
\texttt{qfrac\_eq\_div}: in $\Z$, $a\cdot d = c\cdot b$ with $b, d > 0$
implies $\texttt{qfrac}(a, b) = \texttt{Z2rat}(c)/\texttt{Z2rat}(d)$ in
$\rat$. Denominator positivity is provided by
\texttt{m1\_num\_den\_at\_den\_pos} (product of factorials) and
\texttt{D\_M1\_pos} (\vmcompute).
\end{proof}

\begin{lemma}[$\Mtwo$ paper-form spec equals FLINT entry over $\DMtwo$]\label{lem:M2_spec_eq_int}
\rocq{PrimeGapS1.Cert.M2_spec_eq_int}
\rocqok
\uses{thm:M2_match, thm:M2_spec_rat_eq}
The analogous identity holds for $\Mtwo$:
$\texttt{M2\_spec\_ij}(i, j)
  = \texttt{Z2rat}(M_2^{\mathrm{int}}[i][j]) / \texttt{Z2rat}(\DMtwo)$
for all $(i, j) \in \{0, \ldots, 41\}^2$.
\end{lemma}
\begin{proof}\rocqok
Same shape as Lemma~\ref{lem:M1_spec_eq_int}, with
Theorem~\ref{thm:M2_spec_rat_eq} and Theorem~\ref{thm:M2_match} in
place of their $\Mone$ counterparts. Denominator positivity follows
from \texttt{m2\_num\_den\_at\_den\_pos} (an inductive product of
factorials over the $\texttt{fold\_left qplus}$ accumulator).
\end{proof}

\paragraph{Reading.}
A reviewer who trusts \emph{only Rocq's kernel} obtains, from
Theorem~\ref{thm:M105}, a certified existence of a $\realalg$
eigenvalue of the specific kernel-verified $42 \times 42$ $\Q$-matrix
$M_{105} = 105\cdot\Mone^{-1}\Mtwo$ above $4$ (equivalently, an
eigenvalue of $\Mone^{-1}\Mtwo$ above $4/105$). A reviewer who additionally accepts Maynard's Lemma~8.3
(an Annals-published Rayleigh-quotient identity, paper-side) obtains
$\Mk[105] > 4$ end-to-end.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Trust base}\label{ch:trust}

\section{What is a Qed and what is paper-side}

\begin{itemize}
\item All claims of the form
  $a = b$ in $\Z$ or $\Q$, $\det \ne 0$, root of a polynomial,
  eigenvalue of a $\Q$-matrix, or sign of an algebraic-real polynomial
  evaluation, that appear in the dependency tree of
  Theorem~\ref{thm:M105} are \Qed{} in the source tree.
  Zero \texttt{Axiom}, zero \texttt{Admitted}.
\item The matrix entries themselves (Lemma~8.2 of Maynard) are
  \Qed{} via Theorems~\ref{thm:M1_match} and~\ref{thm:M2_match}: the
  $1764 + 1764$ kernel-checked cross-multiplications close the trust
  loop on the input data.
\item Maynard's Lemma~8.3 ($\Mk = k \cdot \lammax$) is the
  generalised Rayleigh-quotient identity that bridges
  $\lammax > 4/k$ to $\Mk > 4$. This is the analytic step proved in
  the Annals paper (refereed there). It is \emph{not} formalised
  in this repository, by design of the project's scope.
\item The reality of the spectrum of $\Mone^{-1}\Mtwo$ (a consequence
  of $\Mone$ being symmetric positive-definite and $\Mtwo$ symmetric)
  is also paper-side. The Rocq layer extracts a real algebraic
  eigenvalue via IVT (which is automatically real), so it does not
  need to formalise the structural reality claim.
\item The analytic Beta-integral derivation that produces Maynard's
  closed-form $M_1$ and $M_2$ entries (Lemma~8.1 / 8.2 in the paper) is
  taken as a \emph{definition} in Rocq. The kernel certifies that the
  shipped integer matrices match those closed forms, not that the
  closed forms themselves are correct integrations.
\end{itemize}

\section{Reported assumptions}

\texttt{Print Assumptions maynard\_M105\_certified\_pencil} reports
only the Rocq~9.0/9.1 kernel primitives for native 63-bit integers
and \texttt{BigN}: things like $\texttt{Uint63.add}$,
$\texttt{Uint63.mul}$, $\texttt{Uint63.to\_Z}$,
$\texttt{BigN.succ\_spec}$.  These are not project-specific axioms;
they are Rocq's declaration that the OCaml-implemented
native-integer kernel primitives implement their stated
specifications.  Any proof that uses \vmcompute{} on $\Uintsf$
exposes the same set.

The project does not invoke \texttt{native\_compute} anywhere — a
\texttt{grep -n native\_compute theories/S1} returns zero hits in
real proof code.  This keeps the trust-base enumeration minimal.

\section{The FLINT layer is outside the trust base}

The FLINT pipeline (Python + \texttt{python-flint}) is the candidate
generator for the certificate data and an independent cross-check,
but the Rocq proof never invokes Python and never loads the JSON
certificate directly.  The autogenerated witness files are ordinary
Rocq files; if the FLINT layer shipped wrong data, one of the
following \vmcompute{}-based checks would fail:
\begin{itemize}
\item Theorems~\ref{thm:M1_match} and~\ref{thm:M2_match}: the
  $42 \times 42$ matrix entries match Maynard's closed form
  (Lemma~8.2).
\item Lemma~\ref{lem:pencil_clean_match}
  (\texttt{all\_pencil\_clean\_match\_true}): the shipped
  $\texttt{pencil\_int\_clean}$ matrix equals
  $4 D_{M_2} M_1^{\mathrm{int}} - 105 D_{M_1} M_2^{\mathrm{int}}$
  per-cell after the appropriate scaling.
\item \texttt{check\_M1\_det\_710\_true} and
  \texttt{check\_pencil\_det\_710\_true}: the shipped constant-term
  literals for the two integer determinants are correct modulo each
  of $710$ primes.
\item $\texttt{crt\_bound\_M1\_sufficient\_literal}$ and
  $\texttt{crt\_bound\_pencil\_sufficient\_literal}$:
  $2 \cdot \texttt{bound} + 2 \cdot |\texttt{literal}|
   < \texttt{crt\_product\_710}$ for each determinant.
\end{itemize}
Each is a \Qed{} closed by pure kernel arithmetic.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Numerical highlights}\label{ch:numeric}

\paragraph{Matrix dimension.} $42 \times 42$.
Maynard's $\Mk[105]$ construction uses $42$ basis polynomials
$\{F_{b,c} : b + 2c \le 11\}$.

\paragraph{Determinants.}
$\det(M_1^{\mathrm{int}})$ is $2044$ bits, positive;
$\det(\texttt{pencil\_int\_clean})$ is $2613$ bits, negative.

\paragraph{CRT primes.} $710$ \Uintsf\ primes in $[2^{30}, 2^{31})$,
with product exceeding $2^{21300}$ — well above the Hadamard-style
coefficient bounds (${\sim}2^{2044}$ for $M_1^{\mathrm{int}}$,
${\sim}2^{5830}$ for the clean pencil).

\paragraph{Denominators.} $\DMone$ has ${\sim}221$ decimal digits,
$\DMtwo$ ${\sim}227$, and $D_{\mathrm{pencil}}^{\mathrm{clean}}$
${\sim}210$ decimal digits.  All three are strictly positive.

\paragraph{Cauchy bound.} The IVT step uses MathComp's
\texttt{cauchy\_bound} — an explicit rational upper bound on all
real roots derived from the polynomial coefficients.  No custom
bound is introduced.

\paragraph{Build profile.}
On a 16~GB / 6-thread machine: ${\sim}30$--$50$~min wall-clock with
\texttt{make -j6} (clean rebuild).  The dominant single-\Qed{} costs
are:
\begin{itemize}
\item Theorem~\ref{thm:M2_match}
  (\texttt{all\_match\_M2Z\_true}): ${\sim}35$~min sequentially across
  the six $7$-row chunks (\texttt{MaynardVerify/M2\_0..5.v}), reduced
  to a per-chunk wall of a few minutes under \texttt{make -j} —
  $1764$ entries, each up to $36$ rational summands with intermediate
  denominators reaching $\sim 10^{7000}$ digits.
\item Lemma~\ref{lem:crt_primes_all_prime}
  (\texttt{crt\_primes\_710\_all\_prime}): ${\sim}7$~min — $710$
  pure-$\Z$ trial-division primality checks.
\item \texttt{check\_M1\_det\_710\_true},
  \texttt{check\_pencil\_det\_710\_true}: ${\sim}5$--$10$~min each —
  $710 \times 42$-step modular FL plus the constant-term extraction.
\item \texttt{crt\_bound\_M1\_sufficient\_literal} and
  \texttt{crt\_bound\_pencil\_sufficient\_literal}: a few minutes
  each (\vmcompute{} on a few-thousand-digit arithmetic comparison).
\item Theorem~\ref{thm:M1_match}: ${\sim}90$~s.
\item \texttt{Witness.v} / \texttt{Witness\_PencilClean.v} parsing:
  ${\sim}30$--$60$~s.
\end{itemize}

\paragraph{Project size.}  $41$ \texttt{.v} files (33 top-level
files plus $8$ chunk files: \texttt{MaynardVerify/Def.v},
\texttt{MaynardVerify/M2\_0..5.v}, \texttt{CharPolyAgree/Def.v}),
${\sim}15{,}964$ lines total under \texttt{theories/S1/};
\texttt{Witness.v} alone is ${\sim}5{,}700$ lines and
\texttt{Witness\_PencilClean.v} ${\sim}1{,}890$ lines of
autogenerated certificate data.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Map of key lemmas and files}\label{ch:map}

The dependency graph has two independent backbones merging at
\texttt{CertPencil.v}:

\paragraph{Pencil backbone.}
\texttt{Witness} / \texttt{Witness\_PencilClean} $\to$
\texttt{IntMat} / \texttt{IntPoly} $\to$ \texttt{CharPoly} $\to$
\texttt{ModularArith} $\to$ \texttt{CharPolyAgree/Def} +
\texttt{CRTBridge} + \texttt{CRTLift} +
\texttt{Fermat} + \texttt{PrimeCheck} + \texttt{CRTCheck} $\to$
\texttt{CRTPencilCheck} + \texttt{CRTPencilChecksProof} +
\texttt{CRTPencilHadamardGeneric} + \texttt{CRTPencilM1Bound} +
\texttt{CRTPencilPencilBound} + \texttt{PencilCleanGrid} +
\texttt{AbstractPencilHelper} + \texttt{CertPencilDef} +
\texttt{DetPencil} + \texttt{CertL2}.
Headline output: Theorem~\ref{thm:det_int_eq} and
Theorem~\ref{thm:charpoly_neg}.

\paragraph{Maynard-spec backbone.}
\texttt{Witness} $\to$ \texttt{MaynardFactQ} $\to$
\texttt{MaynardBasis} + \texttt{MaynardSpec} $\to$
\texttt{MaynardVerify/} (\texttt{Def.v} + \texttt{M2\_0..5.v} +
assembly). Headline output: Theorems~\ref{thm:M1_match}
and~\ref{thm:M2_match}. The leaf \texttt{MaynardSpecBridge.v} adds
Theorems~\ref{thm:M1_spec_rat_eq} and~\ref{thm:M2_spec_rat_eq} on
top, and \texttt{Cert.v} composes the rat-level matrix identity
\texttt{M\{1,2\}\_spec\_eq\_int} consumed by the headline.

The two backbones merge at \texttt{CertPencil.v}, producing
Theorems~\ref{thm:headline} and~\ref{thm:M105}. A single
\texttt{Require Import PrimeGapS1.CertPencil.} loads the full proof.
